Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-17 methods.tex
introduced data notation in methods.tex
rewrite of ARIMA section and some notation changes there and in methods.tex
added a mention to methods.tex that HPMs can measure other metrics besides IPC

The time-series data for these experiments was collected on an Intel
Core\textsuperscript{\textregistered} i7-2600-based machine running
the 2.6.38-8 Linux kernel.  This particular microprocessor chip has
eight processing units, a clock rate of 3.40GHz, and a cache size of
8192 KB.  We gathered performance traces during the execution of three
different programs---the simple \col loop whose performance is
depicted in Figure~\ref{fig:col-ipc}, as well as two more-complex
programs: one from the SPEC 2006CPU benchmark suite (\gcc), and one
from the LAPACK linear algebra package (\svd).  To record these
measurements, we used the {\tt libpfm4} library, via PAPI
% 
% \footnote{Performance Application Programming Interface}
% 
5.2~\cite{papi}, to stop program execution at 100,000-instruction
intervals and read the contents of the CPU's onboard hardware
performance monitors\footnote{These specialty registers are built into
  modern microprocessors for the purpose of on-board monitoring and
  storage of performance data.}, which we had programmed to count how
many instructions were executed in each clock cycle (IPC).  For an
in-depth description of this custom-measurement infrastructure,
including discussion of the implications of the sampling interval,
please see \cite{zach-IDA10,mytkowicz09,todd-phd}.  For statistical
validation, we collected 15 performance traces from each of the three
programs.  These traces, and the processes that generated them, are
described in more depth in the rest of this section.

>>>>

The time-series data for these experiments was collected on an Intel
Core\textsuperscript{\textregistered} i7-2600-based machine running
the 2.6.38-8 Linux kernel.  This particular microprocessor chip has
eight processing units, a clock rate of 3.40GHz, and a cache size of
8192 KB.  We gathered performance traces during the execution of three
different programs---the simple \col loop whose performance is
depicted in Figure~\ref{fig:col-ipc}, as well as two more-complex
programs: one from the SPEC 2006CPU benchmark suite (\gcc), and one
from the LAPACK linear algebra package (\svd).  In these experiments,
the scalar observation $x_{i}$ at time $i$ was a measurement of how
well each program was able to exploit the native parellelism of the i7
chip.  To record these measurements, we used the {\tt libpfm4}
library, via PAPI
% 
% \footnote{Performance Application Programming Interface}
% 
5.2~\cite{papi}, to stop program execution at 100,000-instruction
intervals and read the contents of the CPU's onboard hardware
performance monitors\footnote{These specialty registers are built into
  modern microprocessors for the purpose of on-board monitoring and
  storage of performance data.  They can be programmed to measure a
  wide array of processor and memory metrics besides the one used
  here.}, which we had programmed to count how many instructions were
executed in each clock cycle (IPC).  For an in-depth description of
this custom-measurement infrastructure, including discussion of the
implications of the sampling interval, please see
\cite{zach-IDA10,mytkowicz09,todd-phd}.  For statistical validation,
we collected 15 performance traces from each of the three programs.
These traces, and the processes that generated them, are described in
more depth in the rest of this section.
